\documentclass[fleqn,11pt]{article}

\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{fixltx2e}
\usepackage{hyperref}
\usepackage{color}
\usepackage{setspace}
\doublespacing

\setlength{\parindent}{1.4em}

\begin{document}
  \lstset{
    language=Python,
    basicstyle=\small,          % print whole listing small
    keywordstyle=\bfseries,
    identifierstyle=,           % nothing happens
    commentstyle=,              % white comments
    stringstyle=\ttfamily,      % typewriter type for strings
    showstringspaces=false,     % no special string spaces
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    frame=tb,
  }

  \title{Progress Report}
  \author{Wilson Fearn, Andrew Hale, Victor Lazaro, Spencer Seeger}
  \date{March 13, 2018}
  \maketitle

  \begin{itemize}
    \item A description of the problem
      \begin{itemize}
        \item The problem that we are solving is allowing machine learning algorithms to take news 
        articles and identify the topics of them. This is applicable to sites that suggest articles 
        for users to read. For example, Google will suggest articles for users based on articles 
        they have read in the past. By using classification, Google would be able to suggest 
        articles that are of the same topic as what a user has read before.
      \end{itemize}
    \item What machine learning model we are initially trying to learn with
      \begin{itemize}
        \item The machine learning algorithms we are planning on using include Na{\"i}ve Bayes, Subject 
        Vector Machines, and Grid Search. All of these learning algorithms can be applied to text 
        classification. We plan on training on all three of them to see which one gives the best overall 
        accuracy, and then optimizing the best machine learning algorithm.
      \end{itemize}
    \item How and from where we are gathering data
      \begin{itemize}
        \item We will be gathering data from many different news websites, including 
        www.washingtonpost.com and www.cnn.com. To get this data, we will have to manually identify 
        website articles and their classification. After this is done, a simple Python script 
        will fetch the article and store it as ``article text, classification'' in a spreadsheet file. 
        The next section includes an instance of our data.
      \end{itemize}
    \item A description of our data set including:
      \begin{itemize}
        \item Actual example instances, including a reasonable representation (continuous, nominal, etc.) and values for each feature
          \begin{itemize}
            \item
            \begin{tabular}{|cc|}
              \hline
              Article                                                             & Classification\\
              \hline
              “Like the Huawei Mate 10 Pro that we reviewed earlier this week...” & Technology\\
              “President Trump on Thursday doubled down on his idea...”           & Political\\
              \hline
            \end{tabular}
          \end{itemize}
        \item How many instances and features we plan to have in our final data set
          \begin{itemize}
            \item 
            We hope to have 400 instances to train and test on, and the features we will have will 
            differ from what our original dataset contains. This is because we will need to transform the 
            data into a different form before handing it to our chosen machine learning algorithm. The text from 
            the articles will be passed to Sklearn which will use Bucket of Words to obtain word counts 
            for each article. This data will then be used for training the model. Thus, as 
            we gather more articles, the final set of features will change due to the inclusion of new 
            words from different articles. Below is an example of what the the machine learning algorithm 
            will actually use to train.
      
            \medskip
            \medskip
            
            \begin{tabular}{|cccccc|}
              \hline
              Upcoming  & Displays  & We      & ... & Like    & Classification\\
              \hline
              0.0543789 & 0.0234958 & 0.05830 & ... & 0.03093 & Technology\\
              0.0907910 & 0.0000    & 0.1257  & ... & 0.07839 & Political\\
              \hline
            \end{tabular}
          \end{itemize}
      \end{itemize}
      \item Brief discussion of plans and schedule to finish the project
      \begin{itemize}
        \item March 13th, 2018 - Submitted Progress report
        \item March 16th, 2018 - Finalized list of website articles and their classifications
        \item March 19th, 2018 - Finished data processing (grabbing data from web pages)
        \item March 23rd, 2018 - Chose machine learning algorithm to use from original list
        \item March 30th, 2018 - Finished fine tuning of chosen machine learning algorithm
        \item April 6th, 2018 - Completed final draft of project write up and presentation
        \item April 9th, 2018 - Finished final Editing of the project write up and presentation
        \item April 10th, 2018 - Submitted final project
      \end{itemize}
  \end{itemize}
\end{document}